A sample workflow

### 2025-09-26 revision/r2/response_r2.tex

\begin{mybox}
\emph{A couple typos in Box e.g., ``calibrating submodes separately allows to
avoid" missing "us"?}
\end{mybox}

% emwSep22: I changed this in the text and here; your previous version needs an 'it' (makes it possible) but I wanted to remove 'fact' so adjusted more. 
Changed to: ``Calibrating submodels separately makes it easier to avoid the likely reality that [...]''

\begin{mybox}
\emph{\textbf{Supplement Minor comments:}\\
Although the supplement works through an example far simpler than the
types of research questions discussed in the main text, it serves as a
reasonable (and accessible) example of how the workflow could be
implemented. There is a brief discussion towards the end of the
supplement about the ``scalability" of the proposed workflow, but it
leaves something to be desired. Have the authors applied this workflow
(or pieces of it) to examples that are more complex that could be cited,
or is that future work?}
\end{mybox}

We disagree that if the model is not sufficiently complex it leaves `something to be desired.' As we want this workflow to be approachable and leave room for readers to think through how they would advance upon it (to whatever degree of complexity they see as biological relevant), we have left it as is.  % emwSep22:  I actually hate the idea that a model needs to be complex, but it seems common among ecologists reviewing our workflow papers .... I might just say: We disagree that if the model is not sufficiently complex it leaves `something to be desired.' As we want this workflow to be approachable and leave room for readers to think through how they would advance upon it (to whatever degree of complexity they see as biological relevant), we have left it as is.

@@

\begin{mybox}
\emph{Stan model not included so could not be assessed (model1\_nc2.stan), I
recommend sharing for transparency.}
\end{mybox}

Apologies that we forgot to share the Stan code. We have now added the code in the workflow example.

\begin{mybox}
\emph{Figure 15, unclear what the posterior predictive distribution is of,
this detail could be added for transparency. Could also add many
posterior predictive datasets to the same plot and then overlay the
observed data to facilitate easier comparisons between y-tilde and y-obs.}
\end{mybox}

We now specify that the first posterior predictive distribution we show corresponds to the predicted counts across all species and populations, and show 9 draws.
% Hmm I don't understand?
% emwSep22:  I think this must be the figure on pg 15 -- we show just one draw of the PPC (red), we could overlap more (like 10-30) in a lighter shade. And put the observed data black ON TOP so easier to see. I think this would be good to do.... just one draw of a PPC is okay, but not greatI IMHO.

### 2025-09-29 analyses/lpiworkflow.qmd
updating ppc figure

Here we start by visually comparing the posterior predictive distribution of the predicted counts (across all species and populations, red line) to observed data (black dashed line):
```{r}
#| fig-height: 3
#| fig-width: 3
#| echo: false

names <- paste0('logy_pred[', 1:data$N, ']')
retrodictions <- unlist(extract(fit, pars = names))

ggplot() +
  geom_density(data = data.frame(obs = log(data$y)), aes(x = obs), alpha = 0.5, linetype = 'dashed') +
  geom_density(data = data.frame(retrodictions), aes(x = retrodictions), alpha = 0.5, col = 'darkred') +
  theme_bw() +
  labs(y = 'Density', x = 'Counts (log-scale)')

@@

We can also look at some posterior replications separately:

```{r}
#| fig-height: 5
#| fig-width: 5
#| echo: false

n <- sample(1:4000, 9)
retrodictions <- (extract(fit, pars = names))
retrodictions_sub <- sapply(names, function(x){retrodictions[[x]][n]})

ggplot() +
  facet_wrap(~ chain) +
  geom_density(data = data.frame(obs = log(data$y)), aes(x = obs), alpha = 0.5, linetype = 'dashed') +
  geom_density(data = retrodictions_sub_df, aes(x = ret), alpha = 0.5, col = 'darkred') +
  theme_bw() +
  labs(y = 'Density', x = 'Counts (log-scale)') +
  theme(strip.text.x.top = element_blank())

>>>>

Here we start by visually comparing 9 draws from the posterior predictive distribution of counts (across all species and populations, red line) to observed data (black dashed line):
```{r}
#| fig-height: 5
#| fig-width: 5
#| echo: false
#| eval: true

names <- paste0('logy_pred[', 1:data$N, ']')
n <- sample(1:4000, 9)
retrodictions <- (extract(fit, pars = names))
retrodictions_sub <- sapply(names, function(x){retrodictions[[x]][n]})

ggplot() +
  facet_wrap(~ chain) +
  geom_density(data = retrodictions_sub_df, aes(x = ret), alpha = 0.5, col = 'darkred') +
  geom_density(data = data.frame(obs = log(data$y)), aes(x = obs), alpha = 0.5, linetype = 'dashed') +
  theme_bw() +
  labs(y = 'Density', x = 'Counts (log-scale)') +
  theme(strip.text.x.top = element_blank())

@@

We may also overlay several draws (here, 30) on the same plot:
```{r}
#| fig-height: 3
#| fig-width: 3
#| echo: false
#| eval: true

names <- paste0('logy_pred[', 1:data$N, ']')
retrodictions <- extract(fit, pars = names)
m <- t(sapply(names, function(n) retrodictions[[n]][1:30]))

ggplot() +
  geom_density(data = data.frame(retro = m[,1]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,2]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,3]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,4]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,5]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,6]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,7]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,8]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,9]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,10]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,11]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,12]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,13]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,14]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,15]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,16]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,17]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,18]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,19]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,20]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,21]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,22]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,23]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,24]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,25]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,26]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,27]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,28]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,29]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(retro = m[,30]), aes(x = retro), alpha = 0.2, col = 'darkred', linewidth = .1) +
  geom_density(data = data.frame(obs = log(data$y)), aes(x = obs), alpha = 0.2, linetype = 'solid', color = 'white', linewidth = 1.5) +
  geom_density(data = data.frame(obs = log(data$y)), aes(x = obs), alpha = 1, linetype = 'solid') +
  theme_bw() +
  labs(y = 'Density', x = 'Counts (log-scale)')

### 2025-09-29 analyses/lpiworkflow.tex
updating ppc figure

\begin{tcolorbox}[enhanced jigsaw, left=2mm, toprule=.15mm, leftrule=.75mm, rightrule=.15mm, breakable, colframe=quarto-callout-tip-color-frame, colback=white, opacityback=0, arc=.35mm, bottomrule=.15mm]

@@

Here we start by visually comparing the posterior predictive
distribution of the predicted counts (across all species and
populations, red line) to observed data (black dashed line):

@@

We can also look at some posterior replications separately:

>>>>

\begin{tcolorbox}[enhanced jigsaw, opacityback=0, toprule=.15mm, left=2mm, leftrule=.75mm, colframe=quarto-callout-tip-color-frame, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, colback=white, breakable]

@@

Here we start by visually comparing 9 draws from the posterior
predictive distribution of counts (across all species and populations,
red line) to observed data (black dashed line):

@@

We may also overlay several draws (here, 30) on the same plot:
